\documentclass[english]{article}
\usepackage[T1]{fontenc}
%\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=1cm,bmargin=2cm,lmargin=1cm,rmargin=1cm}
\setlength{\parindent}{0bp}
\usepackage{amsmath}
\usepackage{amssymb}
\PassOptionsToPackage{normalem}{ulem}
\usepackage{ulem}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\hypersetup{
	colorlinks = True,
	linkcolor = blue,
	anchorcolor = blue,
	citecolor = blue,
	filecolor = blue,
	urlcolor = blue
}
\usepackage{amsthm}

\newtheorem{definition}{Definition}

\renewcommand\thesubsection{\Alph{subsection}}
\makeatletter
\date{}

\makeatother

\usepackage{babel}
\begin{document}
\title{MAE 243 Written \#1}
\author{Benjamin Pierce \\ \href{mailto:bepierce@ucsd.edu}{bepierce@ucsd.edu}}
\date{\today} 
\maketitle

\section{Problem 1}
\begin{enumerate}
\item State whether the problem is convex or not. Justify your answer
\item State whether you should use a linear programming solver, a mixed integer linear programming
solver, or a nonlinear solver
\item Assume that $c_1, c_2, d_1, d_2, d_3 > 0$
\end{enumerate}
\subsection{}
    \begin{align*}
    & \min_x c^T x \\ 
    & s.t. \ \ Ax > b \\
     & x_i \in [0, 1], \forall i = 1,...,n
    \end{align*}
    This problem is convex because both the objective function and the feasible region are themselves convex. Since the values of the decision variable $x$ are $[0,1]$, a mixed integer LP solver shoud be used.

\subsection{} % TODO
    \begin{align*}
    & \min_{x, y} c_1 x + c_2 y\\ 
    & s.t. \ \ d_1 x^2 + d_2 y^2 \leq d_3 \\
     & x,y \geq 0
    \end{align*} 
    This problem is convex (as both the objective and constraints are convex functions), but since the constraints include a squared term, it's a quadratic programming problem, which you'd need a nonlinear solver for.
    
 \subsection{}
     \begin{align*}
     & \min_{x, y} c_1 x + c_2 xy\\ 
     & s.t. \ \ d_1 x + d_2 y \leq d_3 \\
      & x,y \geq 0
     \end{align*} 
    Due to the $c_2 x y$ term, the objective function is nonlinear and nonconvex as the Hessian of $c_1x + c_2 xy$ is 
    $\big(\begin{smallmatrix}
      0 & c_2\\
      c_2 & 0
    \end{smallmatrix}\big)$, which is \textbf{not} positive semi-definite.
    You would therefore have to use a nonlinear solver. 

\section{Problem 2}
Consider the optimization problem:
$$
\begin{array}{ll}
\max _{x, y} & 10 x+5 y \\
\text { s.t. } & 2 x+5 y \leq 50, \\
& 5 x+y \leq 30 \\
& 8 \geq x \geq 2, \\
& y \geq 0
\end{array}
$$
\subsection{}
The model gives the output as following:
\begin{verbatim}
Objective: 84.78260869565219
Value of x: 4.347826086956522
Value of y: 8.26086956521739
\end{verbatim}
\subsection{}
%To convert a LP to standard form, we need to ensure the following:
%\begin{enumerate}
%\item All variables have non-negativity constraints
%\item All other constraints are expressed as equalities
%\item The right-hand side of each constraint is non-negative.
%\item The problem is a minimization
%\end{enumerate}
We can convert this to standard form with slack variables $s_i$:
$$
\begin{aligned}
\min _{x, y} & -(10 x+5 y) \\
\text { s.t. } & 2 x+5 y+s_1=50 \\
& 5 x+y+s_2=30 \\
& x + s_3=8\\
& x - s_4 =2 \\
& x, y, s_1, s_2, s_3, s_4 \geq 0
\end{aligned}
$$
This can then be written in matrix notation: 
$$
\begin{aligned}
\min _{x}\ \ & c^Tx \\
& \text { s.t. } & Ax = b \\
& x \geq 0
\end{aligned}
$$
Where:
$$
\mathbf{A} = \begin{bmatrix}
2 & 5 & 1 & 0 & 0 & 0 \\
5 & 1 & 0 & 1 & 0 & 0 \\
1 & 0 & 0 & 0 & 1 & 0 \\
1 & 0 & 0 & 0 & 0 & -1 \\
\end{bmatrix}
$$
and $b = [50, 30, 8, 2]^T$, $x = [x, y, s_1, s_2, s_3, s_4]^T$.


\subsection{}
Using the matrix $\mathbf{A}$, we see that our slack variables $s_1, s_2,s_3, s_4$ are basic, whereas our decision variables $x, y$ are non-basic, as they have more then one element per column of $\mathbf{A}$ corresponding with them. 

\subsection{}
If we use the duals from JuMP, we find the first two constraints to be binding. 
\begin{verbatim}
2-element Array{ConstraintRef,1}:
 c1 : 2 x + 5 y ≤ 50.0
 c2 : 5 x + y ≤ 30.0
\end{verbatim}
Which is sensible, as $x$ and $y$ are non-basic
\end{document}